Performance of two different ontological and data access methods

### 2014-07-03 cold2014/magnus.tex
use a tilde to avoid breakage
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@725 43ef1190-90fd-4dd7-a769-d0121f02b80e

The application present users with a web interface where they can
input values for any specification they might want to search for.  In
Figure \ref{initialView} we can see how the application initially
looks for a user. Here the user can input values for the different
specifications, for instance fuel type or CO$_2$ emission. After a
user is done choosing specifications he can execute the actual car
model search. The application extracts the data from the form and
comprises it into another format that can easily be used later on. The
next step begins in the search module. Here it starts a thread for
each possible car model to execute each individual search. In the end
the user is presented with a set of valid models to choose from, which
contain the chosen specifications, e.g. the user have searched for a
model with diesel fuel, automatic transmission, four or more gears and
a maximum of six seats.

@@

Before starting the experiments we will present
abbreviations for each factor because their full name made the graphs
totally illegible.

@@

The next step was to create the planning matrix for this
experiment. Table \ref{3factor} shows all the 8 runs for the
experiment with three factors.  After each run the response time were
logged to that particular run. The response time were then used to
check for significant effect when the experiment were evaluated
\begin{wraptable}{l}{0.67\textwidth}
    \begin{tabular}{ | l | l l l |}
    \hline
    {\bf Run} & {\bf Ontology} & {\bf Fuel type} & {\bf Transmission} \\ \hline
	1 & -vso & Diesel & Automatic Gearbox \\ \hline 
	2 & -co & Diesel & Automatic Gearbox \\ \hline 
	3 & -vso & Unleaded Petrol & Automatic Gearbox \\ \hline 
	4 & -co & Unleaded Petrol & Automatic Gearbox \\ \hline 
	5 & -vso & Diesel & Manual Gearbox \\ \hline 
	6 & -co & Diesel & Manual Gearbox \\ \hline 
	7 & -vso & Unleaded Petrol & Manual Gearbox \\ \hline 
	8 & -co & Unleaded Petrol & Manual Gearbox \\ \hline 
    \end{tabular}
    \caption{Planning matrix for experiment with three factors and
      their levels.}\label{3factor}
\end{wraptable}

Like the rest of the experiments we used DanielPlot to evaluate and a
package in R to visualize the result. Figure \ref{3factorGraph}
shows only one significant factor, namely the \textsf{O} factor in the
upper right corner. In Table \ref{3factorEffect}
we can see the time effects from largest to smallest.  We see that the
ontology factor has the largest effect on approximately 2.2 seconds
which means that one of the ontologies overall is 2.2 seconds
faster. By reading the results from the linear model it is the factor
level \textsf{-vso} which has an overall faster response time.

@@

\subsubsection{Experiment with ten factors}
The last experiment we did was a ten factor experiment. The goal here
was to see which factors would be deemed significant in a complete car
model search. With the results from this experiment we could discuss
several aspects of using the different ontologies.  Here we added 4
new factors in comparison to the six factor experiment. The newly
added factors were \emph{weight}, \emph{nr. of doors},
\emph{nr. of seats} and \emph{fuel consumption}. We used the same
approach here as in the previous experiments to calculate the average
value for the new factors. The average values can be seen in the
planning matrix in appendix \ref{appB}. With 10 factors the amount of
runs were $2^{10}$, which made the planning matrix too big to show here.

In the graph \ref{10factorGraph} we can see the results from running
all the 1024 runs.  We can see that there are a lot of significant
effects present here. There are more than any of the previous
experiments, but it is also more factors and factor interactions.
Unfortunately there are so many significant factors that there are
some of them that are illegible, but most of them have a low impact on
the significance graph.  In Table \ref{10factorEffect} we can see the top
twenty absolute effects. We can see that emission also here are present in a lot of the
entries, but so is ontology as well. The new thing here is the fuel
consumption has entered the most significant factors, and is by itself
higher than emission. The linear model again tells us that the option
\textsf{-vso} has a significantly lower response time than
\textsf{-co}

@@

In the Figure \ref{3factorGraph} we can see that only one significant
effects, the O factor. In the results data found in appendix
\ref{appD} we can see that the Configuration Ontology was almost two
seconds slower than the VSO/COO runs.  Three out of four CO runs kept
the response time around two seconds, but one run had to use
approximately 3.5 seconds which resulted in an average run time of
$\approx2.3$ seconds.  In a study done to research the amount of time
a user would be willing to wait, they found out that the tolerable
waiting time for any information retrieval was approximately 2
seconds.~\cite{waitTime} This means that with three factors we were
still within the time frame for a good user experience.

@@

Another factor that would influence the results is the amount of
values the search has to iterate over.  In Table \ref{averages}, we
presented all the average values and the average number of values
present in the Daimler and Renault data. We can see that CO have
$0.54$ more fuel types on average to iterate over and $0.3$ more
transmissions. This is not a lot, but can account for some delay in
the response time.

@@

In Figure \ref{10factorGraph} we can see the results from this
experiment. There are unfortunately a lot of illegible significant
factors, but these are the less significant ones. In Table
\ref{10factorEffect} we see the top twenty absolute effects and
similar to all the other experiments, O is on top of the list. The
average response time of a run done against CO was $\approx58.5$
seconds while a run against VSO/COO had an average response time of
$\approx0.3$ seconds.  This corresponds with the absolute effect of
the O factor to be 58 seconds.

@@

If we take another look at Table \ref{10factorEffect} we see that
there are only three single factors except O that are in the top 15
significant factors.  After the 15 significant factor the absolute
effect drops significantly. The three factors are E, FC and T. E and T
are present because of the same reasons as in the six factor
experiment. We also see in the top 15 that there are only these three
factors present in all the factor interactions as well, which
indicates that T, FC and E affects each other. This is to be expected
because if a car has a high fuel consumption it also has a high
emission rate. There are some possible explanations for why there are
so many significant factor interactions in some of the
experiments. The order will affect this as mention above. Another
explanation is that we chose an average value to be an upper and lower
limit of a factor. We used these as factor levels to avoid an
unrealistic high number of levels. This could have opened for more new
specifications to check on one level of the factor.  As we see in the
Table \ref{10factorEffect}, \textsf{E:WT} is higher than \textsf{WT}
itself. This might indicate that one of the levels of E will open up
for more values to traverse in one of the levels of WT. This is a
complex matter which will depend on each product model as well as the
interaction between the specifications.

>>>>

The application present users with a web interface where they can
input values for any specification they might want to search for.  In
Figure~\ref{initialView} we can see how the application initially
looks for a user. Here the user can input values for the different
specifications, for instance fuel type or CO$_2$ emission. After a
user is done choosing specifications he can execute the actual car
model search. The application extracts the data from the form and
comprises it into another format that can easily be used later on. The
next step begins in the search module. Here it starts a thread for
each possible car model to execute each individual search. In the end
the user is presented with a set of valid models to choose from, which
contain the chosen specifications, e.g. the user have searched for a
model with diesel fuel, automatic transmission, four or more gears and
a maximum of six seats.

@@

For brevity, we will use abbreviations given in Figure~\ref{facandabbrev}.

@@

The next step was to create the planning matrix for this
experiment. Table~\ref{3factor} shows all the 8 runs for the
experiment with three factors.  After each run the response time were
logged to that particular run. The response time were then used to
check for significant effect when the experiment were evaluated
\begin{wraptable}{l}{0.67\textwidth}
    \begin{tabular}{ | l | l l l |}
    \hline
    {\bf Run} & {\bf Ontology} & {\bf Fuel type} & {\bf Transmission} \\ \hline
	1 & -vso & Diesel & Automatic Gearbox \\ \hline 
	2 & -co & Diesel & Automatic Gearbox \\ \hline 
	3 & -vso & Unleaded Petrol & Automatic Gearbox \\ \hline 
	4 & -co & Unleaded Petrol & Automatic Gearbox \\ \hline 
	5 & -vso & Diesel & Manual Gearbox \\ \hline 
	6 & -co & Diesel & Manual Gearbox \\ \hline 
	7 & -vso & Unleaded Petrol & Manual Gearbox \\ \hline 
	8 & -co & Unleaded Petrol & Manual Gearbox \\ \hline 
    \end{tabular}
    \caption{Planning matrix for experiment with three factors and
      their levels.}\label{3factor}
\end{wraptable}

Like the rest of the experiments we used DanielPlot to evaluate and a
package in R to visualize the result. Figure~\ref{3factorGraph}
shows only one significant factor, namely the \textsf{O} factor in the
upper right corner. In Table~\ref{3factorEffect}
we can see the time effects from largest to smallest.  We see that the
ontology factor has the largest effect on approximately 2.2 seconds
which means that one of the ontologies overall is 2.2 seconds
faster. By reading the results from the linear model it is the factor
level \textsf{-vso} which has an overall faster response time.

@@

\subsubsection{Experiment with ten factors}
The last experiment we did was a ten factor experiment. The goal here
was to see which factors would be deemed significant in a complete car
model search. With the results from this experiment we could discuss
several aspects of using the different ontologies.  Here we added 4
new factors in comparison to the six factor experiment. The newly
added factors were \emph{weight}, \emph{nr. of doors},
\emph{nr. of seats} and \emph{fuel consumption}. We used the same
approach here as in the previous experiments to calculate the average
value for the new factors. The average values can be seen in the
planning matrix in appendix~\ref{appB}. With 10 factors the amount of
runs were $2^{10}$, which made the planning matrix too big to show here.

In the graph~\ref{10factorGraph} we can see the results from running
all the 1024 runs.  We can see that there are a lot of significant
effects present here. There are more than any of the previous
experiments, but it is also more factors and factor interactions.
Unfortunately there are so many significant factors that there are
some of them that are illegible, but most of them have a low impact on
the significance graph.  In Table~\ref{10factorEffect} we can see the top
twenty absolute effects. We can see that emission also here are present in a lot of the
entries, but so is ontology as well. The new thing here is the fuel
consumption has entered the most significant factors, and is by itself
higher than emission. The linear model again tells us that the option
\textsf{-vso} has a significantly lower response time than
\textsf{-co}

@@

In the Figure~\ref{3factorGraph} we can see that only one significant
effects, the O factor. In the results data found in appendix
\ref{appD} we can see that the Configuration Ontology was almost two
seconds slower than the VSO/COO runs.  Three out of four CO runs kept
the response time around two seconds, but one run had to use
approximately 3.5 seconds which resulted in an average run time of
$\approx2.3$ seconds.  In a study done to research the amount of time
a user would be willing to wait, they found out that the tolerable
waiting time for any information retrieval was approximately 2
seconds.~\cite{waitTime} This means that with three factors we were
still within the time frame for a good user experience.

@@

Another factor that would influence the results is the amount of
values the search has to iterate over.  In Table~\ref{averages}, we
presented all the average values and the average number of values
present in the Daimler and Renault data. We can see that CO have
$0.54$ more fuel types on average to iterate over and $0.3$ more
transmissions. This is not a lot, but can account for some delay in
the response time.

@@

In Figure~\ref{10factorGraph} we can see the results from this
experiment. There are unfortunately a lot of illegible significant
factors, but these are the less significant ones. In Table
\ref{10factorEffect} we see the top twenty absolute effects and
similar to all the other experiments, O is on top of the list. The
average response time of a run done against CO was $\approx58.5$
seconds while a run against VSO/COO had an average response time of
$\approx0.3$ seconds.  This corresponds with the absolute effect of
the O factor to be 58 seconds.

@@

If we take another look at Table~\ref{10factorEffect} we see that
there are only three single factors except O that are in the top 15
significant factors.  After the 15 significant factor the absolute
effect drops significantly. The three factors are E, FC and T. E and T
are present because of the same reasons as in the six factor
experiment. We also see in the top 15 that there are only these three
factors present in all the factor interactions as well, which
indicates that T, FC and E affects each other. This is to be expected
because if a car has a high fuel consumption it also has a high
emission rate. There are some possible explanations for why there are
so many significant factor interactions in some of the
experiments. The order will affect this as mention above. Another
explanation is that we chose an average value to be an upper and lower
limit of a factor. We used these as factor levels to avoid an
unrealistic high number of levels. This could have opened for more new
specifications to check on one level of the factor.  As we see in the
Table~\ref{10factorEffect}, \textsf{E:WT} is higher than \textsf{WT}
itself. This might indicate that one of the levels of E will open up
for more values to traverse in one of the levels of WT. This is a
complex matter which will depend on each product model as well as the
interaction between the specifications.
